---
title: "Class 10"
format: html
---

1. Importing Data
```{r}
candy_file <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"
candy = read.csv(candy_file)
head(candy)
```

Q1. How many different candy types are in this dataset?
A1: 9 types

Q2. How many fruity candy types are in the dataset?
A2: 37 types

2. What is your favorate candy?
```{r}
candy["Kit Kat", ]$winpercent

```

Q3. What is your favorite candy in the dataset and what is it’s winpercent value?
A3: Almond Joy Winpercent: 50.347545

Q4. What is the winpercent value for “Kit Kat”?
A4: 76.7686

Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?
A5: 49.653503

SKIMR
```{r}
library("skimr")
skim(candy)
```
Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?
A6: Pluribus average values significantly greater

Q7. What do you think a zero and one represent for the candy$chocolate column?
A7: A "1" means "yes" the candy is chocolate, "0" means "No"

Q8. Plot a histogram of winpercent values
```{r}
m <- ggplot(candy_file, aes(x = winpercent)) +
    geom_histogram()
```
Q9. Is the distribution of winpercent values symmetrical?
A9: No
Q10. Is the center of the distribution above or below 50%?
A10: Above
Q11. On average is chocolate candy higher or lower ranked than fruit candy?
A11: Higher
Q12. Is this difference statistically significant?

3. Overall Candy Rankings
```{r}
tail(candy[order(candy$winpercent),], n=5)
```
Q13. What are the five least liked candy types in this set?
A13: Least to most: Nik L Nip, Chiclets, Super Bubble, and Jawbusters
Q14. What are the top 5 all time favorite candy types out of this set?
Q14: Low to high: Snickers, Kit Kat, Twix, Reese's Miniatures, Reese's Peanut Butter Cup

Q15. Make a first barplot of candy ranking based on winpercent values.
```{r}
library()

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```
Q17. What is the worst ranked chocolate candy?
A17: Sixlets
Q18. What is the best ranked fruity candy?
A18: Starbursts

4. Taking a look at pricepercent
```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```
Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?
A19: Fruity Candy
Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?
A20: Most expensive: Nik n Lip, Nestle Smarties, Ring pop, Hershey's Krackel, Hershey's Milk Chocolate. Nik n Lip is least popular
```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

5 Exploring the correlation structure
```{r}
install.packages(corrplot)
library(corrplot)
```

Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?
A22: Anti is chocolatey and fruity candy
Q23. Similarly, what two variables are most positively correlated?
A23: Most positively correlated is Nougat & Bar

6. Principal Component Analysis
```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?
A24: Fruity and Hard are picked up strongly in the positive direction. Generally fruity candies are harder to chew.
