---
title: "Lab 19: Pertussis Mini Project"
author: "Forrest Wang A15911047"
format: pdf
---

#1 Investigating pertussis cases by year

```{r}
library("datapasta")
```

```{r}
library(ggplot2)
```


```{r}
cdc <- data.frame(
                 V1 = c(1922L,1923L,1924L,1925L,
                        1926L,1927L,1928L,1929L,1930L,1931L,1932L,1933L,
                        1934L,1935L,1936L,1937L,1938L,1939L,1940L,1941L,
                        1942L,1943L,1944L,1945L,1946L,1947L,1948L,1949L,
                        1950L,1951L,1952L,1953L,1954L,1955L,1956L,1957L,
                        1958L,1959L,1960L,1961L,1962L,1963L,1964L,1965L,
                        1966L,1967L,1968L,1969L,1970L,1971L,1972L,1973L,
                        1974L,1975L,1976L,1977L,1978L,1979L,1980L,1981L,
                        1982L,1983L,1984L,1985L,1986L,1987L,1988L,1989L,
                        1990L,1991L,1992L,1993L,1994L,1995L,1996L,1997L,
                        1998L,1999L,2000L,2001L,2002L,2003L,2004L,2005L,
                        2006L,2007L,2008L,2009L,2010L,2011L,2012L,2013L,
                        2014L,2015L,2016L,2017L,2018L,2019L),
                 V2 = c(107463L,164191L,165418L,
                        152003L,202210L,181411L,161799L,197371L,166914L,
                        172559L,215343L,179135L,265269L,180518L,147237L,
                        214652L,227319L,103188L,183866L,222202L,191383L,191890L,
                        109873L,133792L,109860L,156517L,74715L,69479L,
                        120718L,68687L,45030L,37129L,60886L,62786L,31732L,
                        28295L,32148L,40005L,14809L,11468L,17749L,17135L,
                        13005L,6799L,7717L,9718L,4810L,3285L,4249L,3036L,
                        3287L,1759L,2402L,1738L,1010L,2177L,2063L,1623L,
                        1730L,1248L,1895L,2463L,2276L,3589L,4195L,2823L,
                        3450L,4157L,4570L,2719L,4083L,6586L,4617L,5137L,
                        7796L,6564L,7405L,7298L,7867L,7580L,9771L,11647L,
                        25827L,25616L,15632L,10454L,13278L,16858L,27550L,
                        18719L,48277L,28639L,32971L,20762L,17972L,18975L,
                        15609L,18617L)
       )
```


Q1. With the help of the R “addin” package datapasta assign the CDC pertussis case number data to a data frame called cdc and use ggplot to make a plot of cases numbers over time.
```{r}
ggplot(cdc) +
  aes(x = V1, y = V2) +
  geom_point() +
  geom_line() +
  ggtitle("Pertussis Cases by Year (1922-2019") +
  xlab("Years") + 
  ylab("Number of Cases")
```



#2. A tale of two vaccines (wP & aP)



Q2. Using the ggplot geom_vline() function add lines to your previous plot for the 1946 introduction of the wP vaccine and the 1996 switch to aP vaccine (see example in the hint below). What do you notice?

```{r}
ggplot(cdc) +
  aes(x = V1, y = V2) +
  geom_point() +
  geom_line() +
  ggtitle("Pertussis Cases by Year (1922-2019") +
  xlab("Years") + 
  ylab("Number of Cases") + 
  geom_vline(xintercept=1946, linetype="dashed", 
                color = "blue", size=1)+
  geom_vline(xintercept=1966, linetype="dashed", 
                color = "red", size=1)+
  geom_text(aes(x=1946, label="wP", y=2e+05), colour="blue", angle=90, vjust = 1.2, text=element_text(size=20))+
  geom_text(aes(x=1966, label="aP", y=2e+05), colour="red", angle=90, vjust = -1, text=element_text(size=20))

```
Q3. Describe what happened after the introduction of the aP vaccine? Do you have a possible explanation for the observed trend?
A3. Pertussis cases per year were at a steady decline following wP introducction. After 1966 with aP introduction, cases were very low, below 5,000 per year up until 1993. A possible explanation for this period of low cases are due to the fact that aP used PURE ANTIGENS and with LESS SIDE EFFECTS, encouraging more vaccine use.

HOWEVER, it's RESURGANCE may be caused by bacterial immunity overtime to the aP vaccine and vaccine hessitancy 



#3. Exploring CMI-PB data
```{r}
library(jsonlite)
```

```{r}
subject <- read_json("https://www.cmi-pb.org/api/subject", simplifyVector = TRUE) 
head(subject, 3)
```

Q4. How may aP and wP infancy vaccinated subjects are in the dataset?
```{r}
table(subject$infancy_vac)
```
A4. There are a total of 96 infancy vaccinated subjects in the dataset, 47 of which received wP, and 49 aP

Q5. How many Male and Female subjects/patients are in the dataset?
```{r}
table(subject$biological_sex)
```
A5: There are 30 male subjects and 66 female subjects

Q6. What is the breakdown of race and biological sex (e.g. number of Asian females, White males etc…)?
A6:
```{r}
table(subject$race, subject$biological_sex)
```

#Side-Note: Working with dates

```{r}
library(lubridate)
time_length( today() - ymd("2000-01-01"),  "years")
```

Q7. Using this approach determine (i) the average age of wP individuals, (ii) the average age of aP individuals; and (iii) are they significantly different?
A7:
(i) & (ii)
```{r}
dates <- subject$year_of_birth
subject$age <- today() - ymd(dates)
subject$age
```

```{r}
library(dplyr)

ap <- subject %>% filter(infancy_vac == "aP")
round( summary( time_length( ap$age, "years" ) ) )

# wP
wp <- subject %>% filter(infancy_vac == "wP")
round( summary( time_length( wp$age, "years" ) ) )
```

(iii)
The average age of aP is 25 while wP is 36

Q8. Determine the age of all individuals at time of boost?
A9:
```{r}
int <- ymd(subject$date_of_boost) - ymd(subject$year_of_birth)
age_at_boost <- time_length(int, "year")
head(age_at_boost)
int/365
```

Q9. With the help of a faceted boxplot (see below), do you think these two groups are significantly different?
A9. According to the boxplot below, and because p-value is less than 0.5, there is a significance difference between the 2 groups
```{r}
ggplot(subject) +
  aes(time_length(age, "year"),
      fill=as.factor(infancy_vac)) +
  geom_histogram(show.legend=FALSE) +
  facet_wrap(vars(infancy_vac), nrow=2) 
```
```{r}
x <- t.test(time_length( wp$age, "years" ),
       time_length( ap$age, "years" ))

x$p.value
```



#Joining multiple tables
```{r}
specimen <- read_json("https://www.cmi-pb.org/api/specimen", simplifyVector = TRUE) 
titer <- read_json("https://www.cmi-pb.org/api/ab_titer", simplifyVector = TRUE) 
```

Q9. Complete the code to join specimen and subject tables to make a new merged data frame containing all specimen records along with their associated subject details:
```{r}
meta <- full_join(specimen, subject)
dim(meta)
head(meta)
```

Q10. Now using the same procedure join meta with titer data so we can further analyze this data in terms of time of visit aP/wP, male/female etc.

```{r}
abdata <- inner_join(titer, meta)
dim(abdata)
```

Q11. How many specimens (i.e. entries in abdata) do we have for each isotype?
```{r}
table(abdata$isotype)
```
A11. For each isotype (IgE  IgG IgG1 IgG2 IgG3 IgG4), their respetive entries in abdata are: 6698, 1413, 6141, 6141, 6141, & 6141 

Q12. What do you notice about the number of visit 8 specimens compared to other visits?
A12: Compared to the other visits, the number of visit 8 speciments are drastically lower, in the 2 digits as opposed to 4 digits
```{r}
table(abdata$visit)
```



#4. Examine IgG1 Ab titer levels
```{r}
ig1 <- abdata %>% filter(isotype == "IgG1", visit!=8)
head(ig1)
```
Q13. Complete the following code to make a summary boxplot of Ab titer levels for all antigens:
```{r}
ggplot(ig1) +
  aes(MFI, antigen) +
  geom_boxplot() + 
  facet_wrap(vars(visit), nrow=2)
```

Q14. What antigens show differences in the level of IgG1 antibody titers recognizing them over time? Why these and not others?
A14. FIM2/3 shows differences in level of IgG1 antibody titers overtime. Others do not show any difference as their MFI is much lower.







